\subsection{CSAIL Directory}
\label{sec:csail-directory-evaluation}

The CSAIL directory is an online directory of about 1000 faculty, staff and students in the MIT Computer Science and Artificial Intelligence Laboratory\footnote{\url{https://www.csail.mit.edu/peoplesearch}}. Each entry contains a person's name, phone number, office number, email address, and position.

Some data, such as a phone number, may be missing from the directory. Still, we expect our framework to be useful in flagging discrepancies between different records. Since the notion of what constitutes an outlier here is imprecise at best, we also expect the tool to allow the user to explore different sets of parameters. To illustrate the process, we dicuss the results returned by three distinct iterations of the tool in the next section, each with increasingly strict limits on the number of outliers returned. Because the CSAIL test set is exclusively textual, we use the histogram model for evaluation.

\subsubsection{Initial run: low specificity filtering}
The search for outliers is initiated with parameters $\theta = 0.8, \epsilon = 0.2$. Correlation detection is disabled for these experiments.

This invocation produces a long list of outliers; a small subset of these is presented below. For privacy reasons, names,  phone numbers, office numbers, and emails have been omitted or anonymized in the following listings.

\begin{lstlisting}[gobble=2]
  Hacker, Alyssa, 32-D968,
    $aph@CSAIL.MIT.EDU$, Postdoctoral Associate

  Bitdiddle, Ben, $32-221B$,
    bbitdid@csail.mit.edu, Graduate Student

  $Lu-ater$, Eva, 32-G972,
    eva@csail.mit.edu, Research Scientist
\end{lstlisting}

In total, 451 entries contain outliers, out of a total of 1000. Office numbers are often flagged, as well as names and email addresses. By changing the input parameters to $\theta = 0.8, \epsilon = 0.2$, most of the outliers due to office numbers disappear due to the lower sensitivity. \lstinline{Hacker, Alyssa} disappears from the list, since e-mails with the domain in all capitalization occur frequently enough in the database that they are not considered outliers at sensitivity level $\epsilon = 0.05$. In total there are $68$ outliers.


%\begin{lstlisting}[gobble=2]
%  Bitdiddle, Ben, $32-221B$,
%    bbitdid@csail.mit.edu, Graduate Student
%  > Value $'32-221B'$ (3) doesn't match
%    feature $'signature'$
%
%  $Lu-ater$, Eva, 32-G972,
%    eva@csail.mit.edu, Research Scientist
%  > Value $'Lu-ater'$ (1) doesn't match
%    feature $'title case'$
%\end{lstlisting}


In addition to identifying outliers, dBoost is equipped with tools that provide the user with additional feedback on why features were identified as outliers.

\begin{lstlisting}[gobble=2]
  Bitdiddle, Ben, $32-221B$,
    bbitdid@csail.mit.edu, Graduate Student
  > Value $'32-221B'$ (3) doesn't match
    feature $'signature'$
  • histogram for ($'signature'$,):
    ██████████
    Lu,Lu,Nd,Nd,Pd,Nd,Nd,Nd
    Lu,Nd,Nd,Pd,Nd,Nd,Nd
    Nd,Nd,Lu,Pd,Nd,Nd,Nd
    ████████████████████ Nd,Nd,Pd,Lu,Nd,Nd,Nd
    ██ Nd,Nd,Pd,Lu,Nd,Nd,Nd,Lu
    ██████ Nd,Nd,Pd,Nd,Nd,Nd
    $█ Nd,Nd,Pd,Nd,Nd,Nd,Lu$
    Nd,Nd,Pd,Nd,Nd,Nd,Nd
    Nd,Pd,Nd,Nd,Nd

  $Lu-ater$, Eva, 32-G972,
    eva@csail.mit.edu, Research Scientist
  > Value $'Lu-ater'$ (1) doesn't match
    feature $'title case'$
  • histogram for ($'title case'$,):
    $False$
    ████████████████████ True
\end{lstlisting}

Our tool highlights the incorrect field, and prints the corresponding histogram. The bin in which the suspicious value falls is also highlighted. The \lstinline{signature} case is particularly interesting: to extract the signature of a string, our tool replace each character by the name of its Unicode class: uppercase Latin letters are \lstinline{Lu}, numbers are \lstinline{Nd}, and punctuation signs are \lstinline{Pd}; hence the string \lstinline{32-221B} is converted to \lstinline{Nd,Nd,Pd,Nd,Nd,Nd,Lu}, which does not fall in any of the dominant bins (the most frequent case, \lstinline{Nd,Nd,Pd,Lu,Nd,Nd,Nd}, describes office numbers of like \lstinline{32-G804}, the predominant form of office numbering in the Stata Center).

Manual inspection of the results reveal that most of the outliers reported are actually bad inputs. There are, however, a number of false positives, such as:

\begin{lstlisting}[gobble=2]
  $DeFect$, Cy, 32-D597,
    cydf@csail.mit.edu, Graduate Student
  > Value $'DeFect'$ (1) doesn't match
    feature $'title case'$
  • histogram for ($'title case'$,):
    $False$
    ████████████████████ True
\end{lstlisting}

The case of \lstinline{DeFect} is correct, but our tool notes that it does not adhere to the casing standard derived from other tuples, and thus reports it.
