\section{Evaluation}
\label{sec:evaluation}

We implemented dBoost in approximately 2000 lines of \texttt{python3} code. Our code is publicly available on GitHub under version 3 of the GNU Public License~\footnote{\url{https://github.com/cpitclaudel/6.830}}. The program is made of two parts: a library, and a number of data acquisition front-ends (CSV and SQL are currently supported). The library provides functions for each of the phases previously described.

Table~\ref{table:flags} shows the usage options to run the models supported by dBoost. \fxnote{Should we keep this table?}

\begin{table*}
  \renewcommand{\arraystretch}{1.2}
  \setlength\tabcolsep{3\tabcolsep}

  \label{table:flags}
  \caption{dBoost command line usage.}
  \centering
  \begin{tabular} { l | l | p{10cm} }
    \multicolumn{3}{l}{} \\
    \hline
    Flag & Options & Explanation \\
    \hline
    --gaussian & n\_stdev & Report outliers that fall more than n\_stdev standard deviations away from the mean of the data \\
    --mixture & n\_subpops & Use a model of \texttt{n\_subpops} Gaussians \\
         & threshold & Report outliers above threshold percentile \\
    --histogram & peak\_s & Consider only fields with a peaked distribution with peakiness peak\_s \\
         & outlier\_s & Report values that fall in classes with less than outlier\_s percent \\
    --statistical & epsilon & Give hints to the model for correlations with Pearson $R$ coefficient greater than epsilon \\
  \end{tabular}
\end{table*}

This sections presents the results of running our tool on a mix of real and synthetic datasets.

To evaluate of our tool, we used a mix of synthetic and real datasets, succinctly described below:

\begin{itemize}
\item \emph{Synthetic datasets}
  \begin{description}
  \item[Fizz-Buzz] A mixed textual-numerical dataset in which each record contains two entries: a number, and either ``Fizz'' if the number is divisible by 3, ``Buzz'' if the number is divisible by 5, ``FizzBuzz'' if it is divisible by both, and the number itself (as a string) otherwise. Outliers appear when the second column does not respect these rules; this can be a misplaced ``Fizz'', a missing ``Buzz'', or even a totally different string (e.g. ``Woof'').
  \item[Web logins] A series of three non-numeric datasets in which entries record the time and place of connection for different users. Each user has different connection habits, leading to different types of outliers.
  \item[TODO] Continuous dataset
  \end{description}
\item \emph{Real-world datasets}
  \begin{description}
  \item[CSAIL Directory] A publicly-accessible directory of researchers\fxnote{Should we anonymize this?}, in which each record include a first and a last name, a phone number, an office number, an email, and a job title. Outliers are hard to define mathematically in this case, and we instead demonstrate how the ideas exposed in previous sections of the paper come together to allow for efficient detection of unwanted values.
  \item[Intel lab data] A publicly-available numeric dataset of temperature, light, and humidity measurements. Outliers are due mostly to sensor glitches.
  \end{description}
\end{itemize}

These datasets showcase the power of our methods, both in terms of classification power, and in terms of expressiveness and succinctness when adding new rules to the system\footnote{Indeed, the set of rules used for tuple expansion is user-configurable, and new rules can be easily added; thus, specific knowledge about the data can be taught to the system by users, expressing some soft form of data integrity constraints.}.

The following subsections describe each of the test sets and associated results in greater detail.

\input{fizzbuzz-evaluation}
\input{logins-evaluation}
\input{csail-directory-evaluation}
\input{intel-lab-data-evaluation}
%\input{presidential-campaign-evaluation}
%\input{mimic2-evaluation}
